Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2023-10-11 paper/main.tex
Reworded sentence

Given these insights, what form might a representation of the sum total of a
person's knowledge take? First, we might require a means of systematically
describing or representing the nearly infinite set of possible things a person
could know. Second, we might want to account for potential associations between
different concepts. For example, the concepts of ``fish'' and ``water'' might
be associated in the sense that fish live in water. Third, knowledge may have a
critical dependency structure, such that knowing about a particular concept
might require first knowing about a set of other concepts. For example,
understanding the concept of a fish swimming in water first requires
understanding what fish and water \textit{are}. Fourth, as we learn, our
``current state of knowledge'' should change accordingly. Learning new concepts
should both update our characterizations of ``what is known'' and also
unlock any now-satisfied dependencies of those newly learned concepts so that
they are ``tagged'' as available for future learning.
% TODO: is the second half of this paragraph relevant to this paper??

@@

We developed a simple formula (Eqn.~\ref{eqn:prop}) for using a participant's
responses to a small set of multiple-choice questions to estimate how much the
participant ``knows'' about the concept reflected by any arbitrary coordinate,
$x$, in text embedding space (e.g., the content reflected by any moment in a
lecture they had watched; see \nameref{subsec:traces}). Essentially, the
estimated knowledge at coordinate $x$ is given by the weighted average
proportion of quiz questions the participant answered correctly, where the
weights reflect how much each question is ``about'' the content at $x$. When we
apply this approach to estimate the participant's knowledge about the content
presented in each moment of each lecture, we can obtain a detailed timecourse
describing how much ``knowledge'' the participant has about any part of the
lecture. As shown in Figure~\ref{fig:knowledge-timeseries}A and C, we can apply
this approach separately for the questions from each quiz participants took
throughout the experiment. From just a few questions per quiz (see~\nameref{subsec:traces}), we obtain a
high-resolution snapshot (at the time each quiz was taken) of what the
participants knew about any moment's content, from either of the two lectures
they watched (comprising a total of 1,100 samples across the two lectures).

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should carry some
predictive information about whether the participant is likely to answer that
question correctly or incorrectly. We developed a statistical approach to test this claim.
For each question in turn, we used Equation~\ref{eqn:prop} to estimate each
participant's knowledge at the given question's embedding space coordinate,
using all \textit{other} questions that participant answered on the same quiz.
%For each question in turn, for each
%participant, we used Equation~\ref{eqn:prop} to estimate (using all
%\textit{other} questions from the same quiz, from the same participant) the
%participant's knowledge at the held-out question's embedding coordinate.
For
each quiz, we grouped these estimates into two distributions: one for the
estimated knowledge at the coordinates of \textit{correctly} answered
questions, and another for the estimated knowledge at the coordinates of
\textit{incorrectly} answered questions (Fig.~\ref{fig:predictions}). We then
used independent samples $t$-tests to compare the means of these distributions
of estimated knowledge.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first and last lines.
These sliding windows ramped up and down in length at the
beginning and end of each transcript, respectively. In other words, each transcript's first
sliding window covered only its first line, the second
sliding window covered the first two lines, and so on. This ensured that each
line from the transcripts appeared in the same number ($w$) of sliding windows.
%and was equally represented in the topic model's training corpus.
After performing various standard text preprocessing (e.g., normalizing case,
lemmatizing, removing punctuation and stop-words), we treated the text from
each sliding window as a single ``document,'' and combined these documents
across the two videos' windows to create a single training corpus for the topic model.
%To select an appropriate number of topics for the model, we identified the
%minimum $k$ that yielded at least one ``unused'' topic (i.e., in which all words
%in the vocabulary were assigned zero weight) after training. This indicated that the
%number of topics was sufficient to capture the set of latent themes present in the two
%lectures. We found this value to be $k = 15$ topics.
%Supplementary Figure~\topicWordWeights~displays the distribution of weights over words
%in the vocabulary for each discovered topic, and each topic's top-weighted words may be found
%in Supplementary Table~\topics.

>>>>

Given these insights, what form might a representation of the sum total of a
person's knowledge take? First, we might require a means of systematically
describing or representing the nearly infinite set of possible things a person
could know. Second, we might want to account for potential associations between
different concepts. For example, the concepts of ``fish'' and ``water'' might
be associated in the sense that fish live in water. Third, knowledge may have a
critical dependency structure, such that knowing about a particular concept
might require first knowing about a set of other concepts. For example,
understanding the concept of a fish swimming in water first requires
understanding what fish and water \textit{are}. Fourth, as we learn, our
``current state of knowledge'' should change accordingly. Learning new concepts
should both update our characterizations of ``what is known'' and also
unlock any now-satisfied dependencies of those newly learned concepts so that
they are ``tagged'' as available for future learning.

@@

We developed a simple formula (Eqn.~\ref{eqn:prop}) for using a participant's
responses to a small set of multiple-choice questions to estimate how much the
participant ``knows'' about the concept reflected by any arbitrary coordinate,
$x$, in text embedding space (e.g., the content reflected by any moment in a
lecture they had watched; see \nameref{subsec:traces}). Essentially, the
estimated knowledge at coordinate $x$ is given by the weighted average
proportion of quiz questions the participant answered correctly, where the
weights reflect how much each question is ``about'' the content at $x$. When we
apply this approach to estimate the participant's knowledge about the content
presented in each moment of each lecture, we can obtain a detailed timecourse
describing how much ``knowledge'' the participant has about any part of the
lecture. As shown in Figure~\ref{fig:knowledge-timeseries}A and C, we can apply
this approach to the questions from each quiz taken throughout the experiment, separately.

From just a few questions per quiz (see~\nameref{subsec:traces}), we obtain a
high-resolution snapshot (at the time each quiz was taken) of what the
participants knew about any moment's content, from either of the two lectures
they watched (comprising a total of 1,100 samples across the two lectures).

@@

If we are able to accurately estimate a participant's knowledge about the
content tested by a given question, our estimates of their knowledge should carry some
predictive information about whether the participant is likely to answer that
question correctly or incorrectly. We developed a statistical approach to test this claim.
For each question, in turn, we used Equation~\ref{eqn:prop} to estimate each
participant's knowledge at the given question's embedding space coordinate,
using all \textit{other} questions that participant answered on the same quiz.
For each quiz, we grouped these estimates into two distributions: one for the
estimated knowledge at the coordinates of \textit{correctly} answered
questions, and another for the estimated knowledge at the coordinates of
\textit{incorrectly} answered questions (Fig.~\ref{fig:predictions}). We then
used independent samples $t$-tests to compare the means of these distributions
of estimated knowledge.

@@

We fetched these automated transcripts using the
\texttt{youtube-transcript-api} Python package~\citep{Depo18}. The transcripts
consisted of one timestamped line of text for every few seconds (mean: 2.34~s;
standard deviation: 0.83~s) of spoken content in the video (i.e., corresponding
to each individual caption that would appear on-screen if viewing the lecture
via YouTube, and when those lines would appear). We defined a sliding window
length of (up to) $w = 30$ transcript lines, and assigned each window a
timestamp corresponding to the midpoint between the timestamps for its first and last lines.
These sliding windows ramped up and down in length at the
beginning and end of each transcript, respectively. In other words, each transcript's first
sliding window covered only its first line, the second
sliding window covered the first two lines, and so on. This ensured that each
line from the transcripts appeared in the same number ($w$) of sliding windows.
After performing various standard text preprocessing (e.g., normalizing case,
lemmatizing, removing punctuation and stop-words), we treated the text from
each sliding window as a single ``document,'' and combined these documents
across the two videos' windows to create a single training corpus for the topic model.
